Disease Knowledge Transfer for Alzheimer's Variants

### 2018-03-01 paper/paper.tex
re-fitting everything. fixed the scaling problem ... need to check validation results after fitting

\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,calc,ifthen}
\usepackage{float}
%\usepackage{cancel}
\usepackage[table,usenames,dvipsnames]{xcolor} % for coloured cells in tables
\usepackage{tikz}
% Allows us to click on links and references!
\usepackage{hyperref}
\usepackage{url}
\hypersetup{
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black
}
% Nice package for plotting graphs
% See excellent guide:
% http://www.tug.org/TUGboat/tb31-1/tb97wright-pgfplots.pdf
\usetikzlibrary{plotmarks,shapes}
\usepackage{amsmath,graphicx}
\usepackage{mathtools}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
% highlight - useful for TODOs and similar
\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{yellow}{#1}}
\newcommand\ci{\perp\!\!\!\perp} % perpendicular sign
\newcommand*\rfrac[2]{{}^{#1}\!/_{#2}} % diagonal fraction
\newcommand\SLASH{\char`\\}
\usepackage{listings}
% margin size
\usepackage[margin=0.5in]{geometry}
\usepackage{pdfpages}
\usepackage{enumitem} % for nested enumerate numbers 1 1.1 1.1.1

\usepackage{titlesec} % reduce spacing after subsections

% \titlespacing\subsection{0pt}{4pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

@@

\begin{document}
\belowdisplayskip=12pt plus 3pt minus 9pt
\belowdisplayshortskip=7pt plus 3pt minus 4pt

% \rowcolors{2}{gray!25}{white}

% \begin{figure}[H]
% \centering
% \includegraphics[scale=0.5]{../code/synthetic/14/figures/liksTrueFittedParams}
% \caption{14 clear biomk}
% \end{figure}
% 
% \begin{figure}[H]
% \centering
% \includegraphics[scale=0.5]{../code/synthetic/14/figures/liksTrueFittedParams}
% \caption{11 clear + 3 unclear biomk}
% \end{figure}

@@

% \section{Motivation}
% 
% It is currently believed that neurodegenerative diseases target specific brain networks (Seeley, 2009). Moreover, different symptoms assiciated with different syndromes are thus believed to be a result of different underlying brain areas being affected.

% TITLE Disease Knowledge Transfer for Alzheimer's Variants

\section{Abstract}

@@

% 1. Statement of Novelty/Impact *

We present a novel dementias paradigm (Affected Overlap) and technique (DKT) that can be used to predict, for the first time, the population-level progression of biomarkers in a disease for which no such biomarker data is available, by transferring them from a different disease where they are available.

@@

@@

\begin{figure}[h]
 \centering
 \includegraphics[height=11cm]{figures/disease_knowledge_transfer.png}
 \caption{Outline of the proposed framework for joint modelling of multiple diseases. }
\end{figure}

@@

>>>>

\documentclass{llncs}
\usepackage{amsmath,amssymb,calc,ifthen}
\usepackage{float}
%\usepackage{cancel}
\usepackage[table,usenames,dvipsnames]{xcolor} % for coloured cells in tables
\usepackage{tikz}
% Allows us to click on links and references!
\usepackage{hyperref}
\usepackage{url}
\hypersetup{
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black
}
% Nice package for plotting graphs
% See excellent guide:
% http://www.tug.org/TUGboat/tb31-1/tb97wright-pgfplots.pdf
\usetikzlibrary{plotmarks,shapes}
\usepackage{amsmath,graphicx}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
% highlight - useful for TODOs and similar
\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{yellow}{#1}}
\newcommand\ci{\perp\!\!\!\perp} % perpendicular sign
\newcommand*\rfrac[2]{{}^{#1}\!/_{#2}} % diagonal fraction
\newcommand\SLASH{\char`\\}
\usepackage{listings}
% margin size
\usepackage{pdfpages}
\usepackage{enumitem} % for nested enumerate numbers 1 1.1 1.1.1
% \usepackage{breqn}
% \usepackage[linesnumbered]{algorithm2e}
% \usepackage{algorithmicx,algpseudocode}
% \usepackage{wrapfig} % for allowing text wrapped around the algorithm
% \newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
% \SetCommentSty{mycommfont}

% \usepackage{titlesec}
% \titlespacing*{\section}
% {0pt}{5.5ex plus 1ex minus .2ex}{4.3ex plus .2ex}

@@

\begin{document}

@@

\title{Disease Knowledge Transfer for Alzheimer's Variants}
%
\titlerunning{Disease Knowledge Transfer for Alzheimer's Variants}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{R\u{a}zvan V. Marinescu\inst{1} \and Pere P. Morell\inst{1} \and
Marco Lorenzi\inst{1,4} \and Neil P. Oxtoby\inst{1} \and Alexandra L. Young\inst{1} \and Arman Eshaghi\inst{1,2} \and Timothy J. Shakespeare\inst{3} \and Kier X. X. Yong\inst{3} \and Sebastian J. Crutch\inst{3} \and Daniel C. Alexander\inst{1}, for 
the  Alzheimer’s  Disease  Neuroimaging Initiative\thanks{Data used in preparation of this article were obtained from the Alzheimer's Disease Neuroimaging Initiative (ADNI) database (adni.loni.usc.edu). As such, the investigators within the ADNI contributed to the design and implementation of ADNI and/or provided data but did not participate in analysis or writing of this report. A complete listing of ADNI investigators can be found at: http://adni.loni.usc.edu/wp-content/uploads/how\_to\_apply/ADNI\_Acknowledgement\_List.pdf}}

%
\authorrunning{R\u{a}zvan V. Marinescu et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
% \tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
% Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{Centre for Medical Image Computing, Computer Science Department, University College London, UK
\and 
Queen Square MS Centre, UCL Institute of Neurology, London
\and 
Dementia Research Centre, UCL Institute of Neurology, University College London, UK
\and
University of C\^{o}te d'Azur, Inria Sophia Antipolis, Asclepios Research Project
% Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
% F-91405 Orsay Cedex, France}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
% The abstract should summarize the contents of the paper
% using at least 70 and at most 150 words. It will be set in 9-point
% font size and be inset 1.0 cm from the right and left margins.
% There will be two blank lines before and after the Abstract. \dots

%5. achievement: we built a voxelwise disease progression model that uses sigmoidal traj
%6. model groups together vertices into clusters and builds a common trajectory for all vertices in the cluster
%7. Simulations show bla bla ...
%8. Tested the model on ADNI and stages correlate well with cognitive tests.
%9. Compared the patterns of typical and atypical AD with DRC data

% The model highlights, for the first time, groups of brain vertices that exhibit a similar temporal trajectory over the population. This provides a new way to parcellate the brain that is specific to the temporal trajectory of a particular disease

%%% MAX NR OF WORDS USED, DO NOT ADD MORE %%%%%%

@@

\keywords{Disease Progression Model, Transfer Learning, Manifold Learning, Alzheimer's Disease, Posterior Cortical Atrophy}
\end{abstract}

%%%%%%%%% NO LONGER THAN 12 pages!! otherwise paper will be instantly rejected %%%%%%%%

\section{Introduction}

@@

% alzheimer's -> not homogeneous, there are many varieties -> PCA -> 
There exist several image-based biomarkers for Alzheimer's disease that can be used to track its progression: AV45 Positron Emission Tomography (PET) measuring amyloid-plaque aggregation, AV1451 PET measuring tau tangle aggregation, Flourodeoxyglucose (FDG) PET measuring brain hypometabolism, Magnetic Resonance Imaging (MRI) measuring structural integrity and Diffusion Tensor Imaging (DTI) measuring connectivity integrity. Measuring the exact evolution of these biomarkers over the disease progression can lead to better disease staging of patients, which helps stratification in clinical trials.

% add another disease example at the end
A hypothetical model of disease progression has been previously published by \ref{jack2010hypothetical}, which proposes that the first biomarkers to become abnormal are measures of amyloid beta aggregation, followed by tau abnormalities, hypometabolism, structural MRI-based measures and finally cognitive decline. While this hypothetical model proposes a modality-specific ordering of biomarkers in typical AD, it has also been observed that within the same modality, biomarker measurements also have a spatial ordering of abnormality that correlated with Braak stages: hippocampal volumes and entorhinal measures become abnormal first, followed by other structures within the temporal lobe, followed by parietal and frontal abnormalities. Moreover, this type of spatial ordering of abnormality has been observed to differ in other Alzheimer's variants. For example, in Posterior Cortical Atrophy, posterior regions such as the occipital lobe and superior parietal regions have been shown to become affected. The network vulnerability theory proposed by \cite{seeley2009neurodegenerative} suggests that damage to specific underlying connectivity networks leads to spatially distinct patterns of neurodegeneration. However, the current understanding is that within the same brain regions,

@@

\begin{figure}[h]
 \centering
 \includegraphics[height=9cm]{figures/disease_knowledge_transfer.png}
 \caption{Outline of the proposed framework for joint modelling of multiple diseases. }
\end{figure}

@@

\section{Acknowledgements}

This work was supported by the EPSRC Centre For Doctoral Training in Medical Imaging with grant EP/L016478/1. AE received a McDonald Fellowship from the Multiple Sclerosis International Federation (MSIF, www.msif.org), and the ECTRIMS - MAGNIMS Fellowship. ALY was supported through EPSRC grant EP/J020990/01. NPO and SG received funding from the EU Horizon 2020 research and innovation programme under grant agreement No 666992. SJC was supported by an Alzheimer’s Research UK Senior Research Fellowship and ESRC/NIHR (ES/L001810/1) and EPSRC (EP/M006093/1) grants. DCA's work on this topic has funding from the EU Horizon 2020 research and innovation programme under grant agreement No 666992, as well as EPSRC grants J020990, M006093 and M020533. Data collection and sharing for this project was funded by the Alzheimer's Disease Neuroimaging Initiative (ADNI) (National Institutes of Health Grant U01 AG024904) and DOD ADNI (Department of Defense award number W81XWH-12-2-0012). The Dementia Research Centre is an ARUK coordination center.

%
% ---- Bibliography ---- 
% USE HARVARD STANDARD

\bibliographystyle{unsrtnat}
\begin{thebibliography}{5}

\bibitem{jack2010hypothetical}
Jack, C.R., Knopman, D.S., Jagust, W.J., Shaw, L.M., Aisen, P.S., Weiner, M.W., Petersen, R.C. and Trojanowski, J.Q., 2010. Hypothetical model of dynamic biomarkers of the Alzheimer's pathological cascade. The Lancet Neurology, 9(1), pp.119-128.

\bibitem{bateman2012clinical}
Bateman, R.J., Xiong, C., Benzinger, T.L., Fagan, A.M., Goate, A., Fox, N.C., Marcus, D.S., Cairns, N.J., Xie, X., Blazey, T.M. and Holtzman, D.M., 2012. Clinical and biomarker changes in dominantly inherited Alzheimer's disease. New England Journal of Medicine, 367(9), pp.795-804.

\bibitem{richberg2015multi}
Schmidt-Richberg, A., Guerrero, R., Ledig, C., Molina-Abril, H., Frangi, A.F., Rueckert, D. and Alzheimers Disease Neuroimaging Initiative, 2015, June. Multi-stage Biomarker Models for Progression Estimation in Alzheimer’s Disease. In International Conference on Information Processing in Medical Imaging (pp. 387-398). Springer International Publishing.

\bibitem{fonteijn2012event}
Fonteijn, H.M., Modat, M., Clarkson, M.J., Barnes, J., Lehmann, M., Hobbs, N.Z., Scahill, R.I., Tabrizi, S.J., Ourselin, S., Fox, N.C. and Alexander, D.C., 2012. An event-based model for disease progression and its application in familial Alzheimer's disease and Huntington's disease. NeuroImage, 60(3), pp.1880-1889.

\bibitem {jedynak2012}
Jedynak, B.M., Lang, A., Liu, B., Katz, E., Zhang, Y., Wyman, B.T., Raunig, D., Jedynak, C.P., Caffo, B., Prince, J.L. and Alzheimer's Disease Neuroimaging Initiative, 2012. A computational neurodegenerative disease progression score: method and results with the Alzheimer's Disease Neuroimaging Initiative cohort. Neuroimage, 63(3), pp.1478-1486.

\bibitem{donohue2014estimating}
Donohue, M.C., Jacqmin-Gadda, H., Le Goff, M., Thomas, R.G., Raman, R., Gamst, A.C., Beckett, L.A., Jack, C.R., Weiner, M.W., Dartigues, J.F. and Aisen, P.S., 2014. Estimating long-term multivariate progression from short-term data. Alzheimer's \& Dementia, 10(5), pp.S400-S410.

\bibitem{schiratti2015mixed}
Schiratti, J.B., Allassonniere, S., Routier, A., Colliot, O., Durrleman, S. and Alzheimers Disease Neuroimaging Initiative, 2015, June. A mixed-effects model with time reparametrization for longitudinal univariate manifold-valued data. In International Conference on Information Processing in Medical Imaging (pp. 564-575). Springer International Publishing.

\bibitem{bilgel2016multivariate}
Bilgel, M., Prince, J.L., Wong, D.F., Resnick, S.M. and Jedynak, B.M., 2016. A multivariate nonlinear mixed effects model for longitudinal image analysis: Application to amyloid imaging. NeuroImage, 134, pp.658-670.

\bibitem{seeley2009neurodegenerative}
Seeley, W.W., Crawford, R.K., Zhou, J., Miller, B.L. and Greicius, M.D., 2009. Neurodegenerative diseases target large-scale human brain networks. Neuron, 62(1), pp.42-52.

% \bibitem{bishop2006}
% Bishop, C.M., 2006. Pattern recognition. Machine Learning, 128.

\bibitem{reuter2012longitudinal}
Reuter, M., Schmansky, N.J., Rosas, H.D. and Fischl, B., 2012. Within-subject template estimation for unbiased longitudinal image analysis. Neuroimage, 61(4), pp.1402-1418.

% \bibitem{reuter2010inverse}
% Reuter, M., Rosas, H.D. and Fischl, B., 2010. Highly accurate inverse consistent registration: a robust approach. Neuroimage, 53(4), pp.1181-1196.

\bibitem{dickerson2009cortical}
Dickerson, B.C., Bakkour, A., Salat, D.H., Feczko, E., Pacheco, J., Greve, D.N., Grodstein, F., Wright, C.I., Blacker, D., Rosas, H.D. and Sperling, R.A., 2009. The cortical signature of Alzheimer's disease: regionally specific cortical thinning relates to symptom severity in very mild to mild AD dementia and is detectable in asymptomatic amyloid-positive individuals. Cerebral cortex, 19(3), pp.497-510.

% \bibitem{thompson2001cortical}
% Thompson, P.M., Mega, M.S., Woods, R.P., Zoumalan, C.I., Lindshield, C.J., Blanton, R.E., Moussai, J., Holmes, C.J., Cummings, J.L. and Toga, A.W., 2001. Cortical change in Alzheimer's disease detected with a disease-specific population-based brain atlas. Cerebral Cortex, 11(1), pp.1-16.

\bibitem{crutch2012posterior}
Crutch, S.J., Lehmann, M., Schott, J.M., Rabinovici, G.D., Rossor, M.N. and Fox, N.C., 2012. Posterior cortical atrophy. The Lancet Neurology, 11(2), pp.170-178.

\bibitem{young2015multiple}
Young, A.L., Oxtoby, N.P., Huang, J., Marinescu, R.V., Daga, P., Cash, D.M., Fox, N.C., Ourselin, S., Schott, J.M., Alexander, D.C. and Alzheimers Disease Neuroimaging Initiative, 2015, June. Multiple Orderings of Events in Disease Progression. In International Conference on Information Processing in Medical Imaging (pp. 711-722). Springer International Publishing.

\end{thebibliography}

\clearpage
